Determinants of Canadian Politicians' Popularity on Wikipedia.

### 2022-04-15 outputs/paper/paper.Rmd
Update data import scripts to include time series Wikipedia page view data
Update paper graphs

```{r setup, include = FALSE}
# Load Packages ----------------------------------------------------------------
library(tidyverse)
library(knitr)
library(dplyr)
library(ggplot2)
library(here)         # For referencing files in directory
library(kableExtra)   # For formatting tables
library(gridExtra)    # For arranging plots
library(bookdown)     # For cross referencing and citing
library(ggrepel)      # For adding point labels to plots
```

@@

>>>>

```{r setup, include = FALSE}
# Load Packages ----------------------------------------------------------------
library(tidyverse)
library(knitr)
library(dplyr)
library(ggplot2)
library(here)         # For referencing files in directory
library(kableExtra)   # For formatting tables
library(gridExtra)    # For arranging plots
library(bookdown)     # For cross referencing and citing
library(ggrepel)      # For adding point labels to plots
library(modelsummary) # For creating summary tables
library(performance)  # For evaluating model performance
library(see)          # For graphing model performance

# Load Data --------------------------------------------------------------------
wiki_data <- read_csv(here("inputs/data/wiki_data_pages.csv"))

# Get the total page views for each politician
wiki_data <-
  wiki_data |>
  rowwise() |>
  mutate(page_views = sum(across(starts_with("views"), na.rm = TRUE)))

# Create Party Factor and Position Factor Variable
wiki_data <-
  wiki_data |>
  mutate(
    party = case_when(
      cp == 1 ~ "CPC",
      liberal == 1 ~ "Liberal",
      green == 1 ~ "Green",
      ndp == 1 ~ "NDP",
      bq == 1 ~ "BQ"
    )
  )
```

\newpage

@@

```{r outliers, echo = FALSE, message = FALSE, warning = FALSE}
# Get the positions of jittered points
pos <- position_jitter(width = 0.1, seed = 1)

# Assign Colors to Each Party
party_colour = 
  c(
    "BQ" = "#003F91",
    "Green" = "#6DA34D",
    "CPC" = "#1B9AAA",
    "NDP" = "#F26419",
    "Liberal" = "#C1292E"
  )

# Plot Outliers
wiki_data |>
  ggplot( aes(x = party, y = page_views, label = if_else(page_views > 500000, name, "")), position = pos) +

    # Plot Points
    geom_point(aes(colour = party), alpha = 0.7, position = pos) +

    # Label Points Belonging to Outliers (pages with > 500,000 views)
    geom_text_repel(
      position = pos,
      box.padding = 1,
      segment.curvature = 0.5,
      segment.ncp = 5,
      segment.angle = 90
    ) +

    # Plot Labels
    xlab("Party") +
    ylab("Page Views") +
    labs(
      title = "Page Views by Political Party",
      subtitle = "Pages with Greater than 500,000 Views are Labelled",
      colour = "Party"
      ) +

    # Set Theme and Color Scheme
    theme_minimal() +
    scale_color_manual(values = party_colour) +
    scale_y_continuous(labels = scales::number, breaks = seq(0, 3000000, 500000))
```

\newpage

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# Plot -> filter PM's and Party Leaders
wiki_data |>
  filter(prime_minister == 0, party_leader == 0) |>
  ggplot(aes(x = party, y = page_views, label = if_else(page_views > 500000, name, "")), position = pos) +

  # Plot Points
  geom_point(aes(colour = party), alpha = 0.7, position = pos) +

  # Label Points Belonging to Prime Ministers
  geom_text_repel(
    position = pos,
    box.padding = 1,
    segment.curvature = 0.5,
    segment.ncp = 5,
    segment.angle = 90
  ) +

  # Plot Labels
  xlab("Party") +
  ylab("Page Views") +
  labs(
    title = "Page Views by Political Party",
    subtitle = "Prime Ministers and Party Leaders are Ommitted",
    colour = "Party"
    ) +

  # Set Theme and Color Scheme
  theme_minimal() +
  scale_color_manual(values = party_colour) +
  scale_y_continuous(labels = scales::number)
```

\newpage

```{r tsleaders, echo = FALSE, message = FALSE, warning = FALSE}
electoral_candidates <- c("Justin Trudeau",	"Erin O'Toole", "Jagmeet Singh", "Annamie Paul", "Yves-François Blanchet")

election_ts <-
  wiki_data |>
  filter(name %in% electoral_candidates)

election_ts <-
  election_ts |>
  select(party, name, starts_with("views")) |>

  # Pivot to plot time series data by party
  pivot_longer(
    cols = starts_with("views"),
    names_to = "month",
    names_prefix = "views_m",
    values_to = "views"
  ) |>

  # Create numeric month variable
  mutate(month = as.numeric(month))

# Assign Colors to Each Party Leader
leader_colour = 
  c(
    "Yves-François Blanchet" = "#003F91",
    "Annamie Paul" = "#6DA34D",
    "Erin O'Toole" = "#1B9AAA",
    "Jagmeet Singh" = "#F26419",
    "Justin Trudeau" = "#C1292E"
  )

election_ts |>
  ggplot(aes(x=month, y=views, label = name, colour = name)) +
  geom_line() +
  theme_minimal() +
  theme(legend.position = "top", legend.text = element_text(size = 7)) +

  geom_text_repel(
    data = election_ts |> filter(name == "Justin Trudeau", month == 9),
    aes(label = "Federal Election on September 20th"),
    colour = "black",
    nudge_x = -2.5,
    nudge_y = -50000,
    segment.curvature = 0.5,
    segment.ncp = 5,
    segment.angle = 90,
    size = 3,
    show.legend = FALSE
  ) +

  # Set Labels
  labs(
    x = "Month",
    y = "Pageviews",
    title = "Federal Election Candidates 2021 Monthly Pageviews",
    colour = ""
  ) +

  # Set Scales
  scale_colour_manual(values = leader_colour) +
  scale_x_discrete(limits = 1:12, labels = month.abb[1:12]) +
  scale_y_continuous(labels = scales::number)
```

\newpage

```{r tsparties, echo = FALSE, message = FALSE, warning = FALSE}
parties_ts <-
  wiki_data |>
  group_by(party) |>
  summarize(across(starts_with("views"), ~sum(.x, na.rm = TRUE))) |>

  # Pivot to plot time series data by party
  pivot_longer(
    cols = starts_with("views"),
    names_to = "month",
    names_prefix = "views_m",
    values_to = "views"
  ) |>

  # Create numeric month variable
  mutate(month = as.numeric(month))

parties_ts |>
  ggplot(aes(x=month, y=views, colour = party)) +
  geom_line() +
  theme_minimal() +
  theme(legend.position = "top") +

  # Set Labels
  labs(
    x = "Month",
    y = "Pageviews",
    title = "Federal Parties 2021 Monthly Pageviews",
    colour = ""
  ) +

  # Set Scales
  scale_colour_manual(values = party_colour) +
  scale_x_discrete(limits = 1:12, labels = month.abb[1:12]) +
  scale_y_continuous(labels = scales::number)
```

\newpage

```{r proportions, echo = FALSE, message = FALSE, warning = FALSE}
# Create Election results table
# Seats from https://www.ourcommons.ca/members/en/party-standings (NOTE: 1 Seat to Independent)
# Vote Share From: https://www.bloomberg.com/graphics/2021-canada-election-live-results/
election_results <-
  tibble(
    seats = c("Liberal" = 159, "CPC" = 119, "BQ" = 32, "NDP" = 25, "Green" = 2, "Other" = 1),
    vote_share = c("Liberal" = 0.326, "CPC" = 0.337, "BQ" = 0.076, "NDP" = 0.178, "Green" = 0.023, "Other" = 0.06),
    party = c("Liberal", "CPC", "BQ", "NDP", "Green", "Other")
  )

party_proportions <-
  wiki_data |>

  # Get page views and number of pages by party
  group_by(party) |>
  summarize(
    page_views = sum(page_views, na.rm = TRUE),
    pages = n()
  ) |>

  # Add actual election results by party
  add_row(party = "Other", page_views = 0, pages = 0) |>
  left_join(y = election_results, by = "party") |>

  # Get percent share for each party
  mutate(
    page_views = page_views/sum(page_views),
    pages = pages/sum(pages),
    seats = seats/338
  ) |>

  # Pivot for easier graphing
  pivot_longer(!party, names_to = "type", values_to = "percent")

party_proportions |>

  # Create Plot
  ggplot(aes(x = percent, y = type, fill = party)) +
  geom_bar(stat = "identity", position = "stack") +
  # Text Labels with help from:
  # https://stackoverflow.com/questions/34903368/how-to-center-stacked-percent-barchart-labels
  geom_text(
    aes(
      label = if_else(percent > 0.05, paste0(round(100*percent),"%"), "") 
      ), 
    size = 3,
    position = position_stack(vjust=0.5),
    colour = "white"
    ) +

  # Set labels
  labs(
    title = "Share of Wikipedia Pages, Page Views, and House Seats by Party",
    subtitle = "Page Views, Page Count, and Election Results from the 2021",
    fill = "",
    x = "Percent Share",
    y = ""
  ) +

  # Set Scale to Percentage
  scale_x_continuous(labels = scales::percent) +
  scale_y_discrete(
    labels = c("vote_share" = "Votes", "seats" = "House Seats", "pages" = "Pages", "page_views" = "Page Views")
    ) +

  # Set Theme and Color Scheme
  theme_minimal() +

  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "top"
  ) +

  guides(fill = guide_legend(nrow = 1, byrow=TRUE)) +
  scale_fill_manual(values = c(party_colour, "Other" = "grey"))
```

\newpage

```{r dataexp, echo = FALSE, message = FALSE, warning = FALSE}
# Create a set of bar charts with descriptive statistics by party
# - avg. page views
# - avg. NON-Position Page Views

# - avg. page views by position -> PM, Leader, Minister, Other

# - gender share per party - different table
# - page views per gender
# - Prime Minister Page Views?

position_views <-
  wiki_data |>
  group_by(position) |>
  summarize(page_views = mean(page_views, na.rm = TRUE))

summary_data <-
  wiki_data |>
  mutate(gender = if_else(gender == "male", 1, 0)) |>
  select(page_views, gender, birthyear, prime_minister, party_leader, minister, party)
```

\newpage

```{r balance, echo = FALSE, message = FALSE, warning = FALSE}
# Create a by party balance table (NOTE: Probably Add this in the Appendix, Flip Sideways for Length)
summary_data <-
  wiki_data |>
  mutate(gender = if_else(gender == "male", 1, 0)) |>
  select(page_views, gender, birthyear, prime_minister, party_leader, minister, party) |>
  rename(
    "Page Views" = page_views,
    "Male (%)" = gender,
    "Birth Year" = birthyear,
    "Prime Minister (%)" = prime_minister,
    "Party Leader (%)" = party_leader,
    "Minister (%)" = minister
  )

datasummary_balance(formula = ~ party, data = summary_data, dinm = FALSE, fmt = "%.3f") |>
  kableExtra::landscape() |>
  kable_styling(font_size = 10)
```

\newpage

```{r models, echo = FALSE, message = FALSE, warning = FALSE}
models <- 
  list(
    "Party" = lm(page_views ~ liberal + cp + bq + green, data = wiki_data),
    "Party and Position" = lm(page_views ~ liberal + cp + bq + green + party_leader + prime_minister + minister, 
                    data = wiki_data),
    "Full" = 
      lm(page_views ~ liberal + cp + bq + green + party_leader + prime_minister + minister + gender + 
           birthyear, data = wiki_data |> mutate(birthyear = birthyear - 1950))
)

coef_map = c(
  "liberal" = "Liberal",
  "cp" = "CPC",
  "bq" = "BQ",
  "green" = "Green",
  "prime_minister" = "Prime Minister",
  "party_leader" = "Party Leader",
  "minister" = "Minster",
  "gender" = "Male",
  "birthyear" = "Birthyear - 1950"
)

notes = "Estimate standard error appears in brackets. Birthyear - 1950, measures the number of years between the politician's year of birth and the year 1950. Liberal, CPC (Conservative Party of Canada), Green, and BQ (Bloc Québécois) are party indicators equal to 1 if the politician is a member of the party, and 0 otherwise. The NDP (New Democratic Party) is excluded from the set of party indicators and is used as the reference group. Prime Minister, Party Leader, and Minister are position indicators equal to 1 if the politician has held that position and 0 otherwise. Male is an indicator equal to 1 if the politician is male and 0 otherwise."

modelsummary(
  models,
  coef_map = coef_map, 
  gof_omit = "^(?!R2|Num)",
  notes = notes,
  threeparttable = TRUE
  )
```

```{r modelsrestricted, echo = FALSE, message = FALSE, warning = FALSE}
models_restricted <- 
  list(
    "Party" = lm(page_views ~ liberal + cp + bq + green, 
                 data = wiki_data |> 
                   filter(!(name %in% electoral_candidates))
                 ),
    "Party and Position" = lm(page_views ~ liberal + cp + bq + green + party_leader + prime_minister + minister, 
                    data = wiki_data |> 
                      filter(!(name %in% electoral_candidates))
                    ),
    "Full" = 
      lm(page_views ~ liberal + cp + bq + green + party_leader + prime_minister + minister + gender + 
           birthyear, 
         data = wiki_data |> 
           mutate(birthyear = birthyear - 1950) |> 
           filter(!(name %in% electoral_candidates))
         )
    )

coef_map = c(
  "liberal" = "Liberal",
  "cp" = "CPC",
  "bq" = "BQ",
  "green" = "Green",
  "prime_minister" = "Prime Minister",
  "party_leader" = "Party Leader",
  "minister" = "Minster",
  "gender" = "Male",
  "birthyear" = "Birthyear - 1950"
)

notes = "Estimate standard error appears in brackets. Birthyear - 1950, measures the number of years between the politician's year of birth and the year 1950. Liberal, CPC (Conservative Party of Canada), Green, and BQ (Bloc Québécois) are party indicators equal to 1 if the politician is a member of the party, and 0 otherwise. The NDP (New Democratic Party) is excluded from the set of party indicators and is used as the reference group. Prime Minister, Party Leader, and Minister are position indicators equal to 1 if the politician has held that position and 0 otherwise. Male is an indicator equal to 1 if the politician is male and 0 otherwise."

modelsummary(
  models_restricted,
  coef_map = coef_map, 
  gof_omit = "^(?!R2|Num)",
  notes = notes,
  threeparttable = TRUE
  )
```

\newpage

```{r performance, echo = FALSE, message = FALSE, warning = FALSE}
# Evaluate the Performance of the full model
performance::check_model(models$Full)
```
